### 2017-05-29 src/5_1_Proposal.tex
check formulas references before finishing

\chapter{Adaptive parallel algorithm for Vector error correction model}
VECM (presented in section \ref{sec:vecm}) has been
broadly used with low frequency time series. We aimed to adapt VECM to be used
in finance with high frequency stream data.

We observed that the number of cointegration relationships varies with the
length of historical data used. Moreover, parameters that increased these
relationships in time led to better forecasting performance. Our proposal,
called an Adaptive VECM (AVECM) is to make a parameters grid search that
maximises the number of cointegration relationships in the near past. To ensure
the search can be executed fast enough, we used a distributed environment.

@@

\section{Introduction}
\label{sec:introduction}
In finance, it is common to find variables with long-run equilibrium
relationships. This is termed cointegration and reflects the idea that some sets
of variables cannot wander too far from each other. Cointegration means that one
or more linear combinations of these variables are stationary even though
individually they are not. Some models, such as the Vector Error Correction
(VECM), see \cite{engle87}, take advantage of this property and describe the
joint behaviour of several cointegrated variables. VECM introduces this long-run
relationship among a set of cointegrated time series as an error correction
term. These time series must be integrated of order 1, denoted I(1), i.e. they
become stationary at their first differences. In finance, I(1) time series are
very common and to introduce cointegration restrictions in models often improves
forecasting, see \cite{duy1998}. Therefore, VECM has been widely adopted in
financial applications, among others: \cite{mukherjee1995}, \cite{seong2013},
\cite{maysami2000} and \cite{arestis2001}. VECM has also been used in pair
trading, see \cite{herlemont2003}, or models with more than two variables, see
for example \cite{mukherjee1995} and \cite{engle2004}).

VECM parameters are obtained using the ordinary least squares (OLS) method,
developed by \cite{golub1980}. Since OLS involves many calculations, the
parameter estimation is computationally expensive when the number of lagged
values and data increases. Moreover, obtaining cointegration vectors is also an
expensive routine. This is a main limitation to use VECM with stream data with
high frequency. \cite{chen2003} addresses the advantage of distributed
processing over conventional rolling window processing.

Our aims were to study if a parallel version of VECM can be used with high
frequency stream data.  Our approach was to determine adaptively the number of
observations and lags of VECM which maximise cointegration relations in the
past.  This search is computationally expensive because the Johansen method, see
\cite{johansen1995}, is required to find cointegration vectors, which is of
order $O(n^3)$. Therefore, our proposal, called Adaptively Vector Error
Correction (AVECM), is to parallelise this search in order to get new parameters
before new data arrives.  Model effectiveness is focused on out-of-sample
forecast rather than in-sample fitting. This criterion allows AVECM prediction
capability to be expressed rather than just explaining data history.  The
forecast capability of our method was measured using MSE and the Theil's
$U$-statistic, see \cite{theil1966}, widely used in economic forecast. Tests
were run using four currency rates: Euro (EUR) to United States Dollar (USD)
(EURUSD), British Pound (GBP) to USD (GBPUSD), USD to Swiss Franc (CHF) (USDCHF)
and USD to Japanese Yen (JPY) (USDJPY) with a 10-second frequency.

\section{Methodology} \label{sec:methodology}
\subsection{Motivation} \label{sec:proposal}

Cointegration vectors can be found applying the Johansen method which uses a
sample of the last historical data. However, VECM assumes cointegration vectors
do not change in time.  In fact, \cite{gregoryETal1996} addresses that the
long-run relationships between the time series  might change due to several
economic factors that can lead to structural breaks in the cointegration
relationship.  In order to show that the number of cointegration vectors depends
on the amount $L$ of historical data and the number of lags $p$ in the VECM, we
used a grid search.  We defined a grid of possible values for $L$ and $p$. $L$
goes throughout $[2,14]$ hours (1 $hour = 360$ data points) with a step size of
4 hours and $p$ throughout $[1,5]$ with step size of 1. The first experiment
consisted in determining the number of cointegration vectors for all
combinations of $L$ and $p$.  We used four forex rates: EURUSD, GBPUSD, USDCHF
and USDJPY with 10-second frequency. Data started at 13:00 GMT of the 13th of
August 2014, when the New York and London financial markets opened.

Figure \ref{fig:hists} shows the distribution of the number of cointegration
vectors given by the Johansen method for different values of $L=[2,6,10,14]$
hours  and $p=1$ . This procedure was carried out by a sliding window of
historical data moving 1000 times.

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=0.9\textwidth]{img/Fig1}
  \caption{Distribution of the number of cointegration vectors using $p=1$ lags.
  Four possible values for windows size $L$ are shown (2, 6, 10 and 14) hours (1
  hour = 360 data points).}
  \label{fig:hists}
\end{figure}

From section \ref{sec:background} we know that $r=0$ means no cointegration and
$r=l$ (we are using four rates, so $l=4$) reveals that no process is I(1) but
stationary.  The interesting cases of cointegration are those where $r$ lies
strictly between $0$ and $4$, i.e. $0<r<4$.

@@

\subsection{Model comparison} \label{sec:random}
We compared our proposal, in terms of performance, with the naive forecast of
the random walk model and ARIMA. It is still difficult to outperform the random
walk model for standard econometric forecasting models  despite its
simplicity, see \cite{lo2011}. The random walk model is defined as:

@@

\subsection{Evaluation methods} \label{sec:evaluation}

Forecast performance was evaluated using two different methods which are
frequently used in finance:
\begin{description}
\item
{\bf MSE},  Mean Square Error measures the distance between forecasts
and the true values and large deviations from the true value have a
large impact due to squaring forecast error.
\begin{equation}\label{eq:MSE}
\text{MSE} = 
\frac{\displaystyle \sum_{t=1}^{N} (\mathbf{y}_t-\hat{\mathbf{y}}_t)^2}{N}
\end{equation}
\item {\bf $U$-statistic}, the Theil's $U$-statistic, presented by
\cite{theil1966}, is a unit free measure obtained as the ratio between the root
MSE (RMSE) of a model and the RMSE of the naive random walk model. A
$U$-statistic less than 1 implies the performance is better than the naive
model.
\end{description}

\section{Experimental results}
\label{sec:results}
\subsection{Data} \label{sec:unitroot}
AVECM tests were carried out using four foreign exchange rates all related to
USD: EURUSD, GBPUSD, USDCHF and USDJPY. This data was collected from
\cite{Dukascopy2014}, a free database which gives access to the Swiss Foreign
Exchange marketplace.

@@

\subsection{Unit root tests} \label{sec:unitroottests}

@@

\subsection{Performance accuracy} \label{sec:performance}

@@

\subsection{Parallel implementation} \label{sec:parallel}

@@

\section{Conclusions}
\label{sec:conclusions}
Cointegration in financial time series has been largely studied and the Johansen
method is commonly used to obtain cointegration relationships.  In practice, it
has been found that cointegration relations change with time.  However,
model-based cointegration such as VECM assumes that cointegration remains
unchanged in time.  We empirically showed that the Johansen method is sensitive
to the number of lags but also to the amount of data considered.

>>>>

\chapter{Fast and adaptive cointegration based model
for forecasting high frequency financial time series}

Cointegration is a long-run property of some non-stationary time series where a
linear combination of those time series is stationary.  This behaviour has been
studied in finance because cointegration restrictions often improve forecasting.
The Vector Error Correction Model (VECM) is a well-known econometric technique
that characterises short-run variations of a set of cointegrated time series
incorporating long-run relationships as an error correction term. VECM has been
broadly used with low frequency time series. We aimed to adapt VECM to be used
in finance with high frequency stream data.

Cointegration relations change in time and therefore VECM parameters must be
updated when new data is available. We studied how forecasting performance is
affected when VECM parameters and the length of historical data used change in
time. We observed that the number of cointegration relationships varies with the
length of historical data used. Moreover, parameters that increased these
relationships in time led to better forecasting performance. Our proposal,
called an Adaptive VECM (AVECM) is to make a parameters grid search that
maximises the number of cointegration relationships in the near past. To ensure
the search can be executed fast enough, we used a distributed environment 
\cite{Arce2017}.

@@

\section{Introduction}
In finance, it is common to find variables with long-run equilibrium
relationships. This is termed cointegration and reflects the idea that some sets
of variables cannot wander too far from each other. Cointegration means that one
or more linear combinations of these variables are stationary even though
individually they are not. Some models, such as the Vector Error Correction
(VECM), see \cite{engle87}, take advantage of this property and describe the
joint behaviour of several cointegrated variables. VECM introduces this long-run
relationship among a set of cointegrated time series as an error correction
term. These time series must be integrated of order 1, denoted I(1), i.e. they
become stationary at their first differences. In finance, I(1) time series are
very common and to introduce cointegration restrictions in models often improves
forecasting, see \cite{duy1998}. Therefore, VECM has been widely adopted in
financial applications, among others: \cite{mukherjee1995}, \cite{seong2013},
\cite{maysami2000} and \cite{arestis2001}. VECM has also been used in pair
trading, see \cite{herlemont2003}, or models with more than two variables, see
for example \cite{mukherjee1995} and \cite{engle2004}.

Cointegration relationships can be found in low and high frequency data of two or
more assets. While cointegration in low frequency data is motivated by a
long-run equilibrium relationship between economic forces, cointegration in high
frequency data has its foundation in statistical arbitrage theory which is very helpful
to detect mean-reverting trades. Information about cointegrated assets in high
frequency data could be used as an input for high frequency trading strategies,
using it as a signal to capture small profits in short term trades, see
\cite{miao2014}. \cite{zhou2001} addressed the benefits of using higher
frequency data to analize cointegration. \cite{rittler2012} also found that
cointegration may differ with different data frequencies and could appear
whith increased data frequency.

The use of VECM with high frequency data is mainly limited by computationally
expensive routines. Firstly, VECM parameters are obtained using the ordinary
least squares (OLS) method, developed by \cite{golub1980}. Since OLS involves
many calculations, the parameter estimation is computationally expensive when
the number of lagged values and data increases. Secondly, obtaining
cointegration vectors is also an expensive routine because the Johansen method 
is required, which is of order $O(n^3)$ (\cite{johansen1995}).
\cite{chen2003} addressed the advantage of distributed processing over
conventional rolling window processing. Therefore, our aim was to study if a
parallel version of VECM can be used with high frequency stream data.

Our approach was to determine, adaptively, the number of observations and lags of
VECM which maximise cointegration relations in the past in a distributed
environment. We called our proposal Adaptive Vector Error Correction (AVECM). 
AVECM parallelises this search of parameters in order to update them
before new data arrives. Model effectiveness is focused on out-of-sample
forecast rather than in-sample fitting. This criterion allows AVECM prediction
capability to be expressed rather than just explaining data history.  The
forecast capability of our method was measured using MSE and the Theil's
$U$-statistic, see \cite{theil1966}, widely used in economic forecast. Tests
were run using four currency rates: Euro (EUR) to United States Dollar (USD)
(EURUSD), British Pound (GBP) to USD (GBPUSD), USD to Swiss Franc (CHF) (USDCHF)
and USD to Japanese Yen (JPY) (USDJPY) with a 10-second frequency.

The AVECM algorithm is presented in section \ref{sec:51methodology}.
In section \ref{sec:51results} we describe the tests carried on to assess the
accuracy and the execution time of AVECM.  This section also includes a
description of the test data.  Section \ref{sec:51conclusions} contains the
conclusions and a discussion of future research.

%\section{Background}
%\label{sec:background}
%
%\subsection{Integration and Cointegration}\label{sec:coint}\  
%Following \cite{johansen1995} we shall say that a stochastic process
%$Y_t$ which satisfies $Y_t-E(Y_t) = \sum_{i=0}^\infty C_i\,\varepsilon_{t-i}$ is
%called $I(0)$, and then we shall write $Y_t\sim I(0)$, whenever
%$\sum_{i=0}^\infty C_i \neq 0$ and $\sum_{i=0}^\infty C_i\,z^i$ converges for
%$z\in\mathbb{C}$ with $|z|<1$.  It is understood that the condition
%$\varepsilon_t\sim iid(0,\sigma^2)$ holds.
%
%A (vector) time series $\mathbf{y}_t$ is said to be {\em integrated of order\/}
%$d$, and then we shall write $\mathbf{y}_t\sim I(d)$, whenever after $d$ times
%(discrete) differentiation an stationary process is
%obtained, see \cite{banerjee1993}; more precisely, whenever
%$(1-L)^d\,\mathbf{y}_t\sim\text{I(0)}$, where $L$ is the usual lag operator:
%$(1-L)\,\mathbf{y}_t = \Delta\mathbf{y}_t = \mathbf{y}_t-\mathbf{y}_{t-1}$ for
%all $t$.  
%
%Note that this definition includes the scalar case as time series of vectors of
%dimension 1; in this scalar case we will write the time series in nonbold
%format.
%
%Let $\mathbf{y}_t^\nu$, $\nu=1,\dots,l$, be a set of $l$ vector time series of
%order $I(1)$.  They are said to be {\em cointegrated\/} if a vector
%$\beta=[\beta(1),\dots,\beta(l)]^\top \in \mathbb{R}^p$ exists, such that the
%time series,
%\begin{equation}
%\mathbf{Z}_t:= 
%\sum_{\nu=1}^l \beta(\nu)\,\mathbf{y}_t^\nu\,\sim\,\text{I(0)}\,.
%\end{equation}
%In other words, a set of $I(1)$ time series is said to be cointegrated if a
%linear combination of them exists, which is I(0).
%
%
%\subsection{Vector Autorregresive Models}\label{sec:varvec}
%
%Vector error correction model (VECM) describe the joint behaviour of a set of
%variables and can be derived from the simple Vector Autoregressive model (VAR)
%presented by \cite{sims1980}.  The VAR($p$) model is a framework describing the
%behaviour of a set of $l$ endogenous and stationary variables as a linear
%combination of their last $p$ values, where $l,p\in\mathbb{N}$.  In our case,
%each one of these $l$ variables is a scalar time series $y_{\lambda,t}$,
%$\lambda=1,\dots,l$, and we represent them all together at time $t$ by the
%vector time series:
%\begin{equation}
%\label{eq:variables}
%\mathbf{y}_t = 
%\begin{bmatrix} y_{1,t} & y_{2,t} & \dots & y_{l,t} \end{bmatrix}^\top.
%\end{equation}
%\noindent
%Notice that the vectors $\mathbf{y}_t$ are assumed to be $l$-dimensional.
%
%The VAR($p$) model describes the behaviour of a dependent variable in terms of
%its own lagged values and the lags of the others variables in the system. The
%model with $p$ lags is formulated as the system:
%\begin{align}
%\label{eq:var}
%\mathbf{y}_t 
%= \boldsymbol{\Phi}_1 \mathbf{y}_{t-1} +
%  \boldsymbol{\Phi}_2 \mathbf{y}_{t-2} + \dots +
%  \boldsymbol{\Phi}_p\mathbf{y}_{t-p} +
%  \mathbf{c} + \boldsymbol{\epsilon}_t \nonumber \\
%t=p+1,\dots,N,
%\end{align}
%\noindent where 
%$\boldsymbol{\Phi}_1, \boldsymbol{\Phi}_2,\dots,\boldsymbol{\Phi}_p$
%are $l\times l$-matrices of real coefficients,
%$\boldsymbol{\epsilon}_{p+1},
% \boldsymbol{\epsilon}_{p+2}, \dots, \boldsymbol{\epsilon}_N$ 
%are error terms, $\mathbf{c}$ is a constant vector and $N$ is the total
%number of samples.
%
%Notice that, regarding our notation of section (\ref{sec:coint}),
%we have here 
%$\mathbf{y}_t^0 = \mathbf{y}_t$,
%$\mathbf{y}_t^\nu = \mathbf{y}_{t-\nu}$ and
%the $\lambda$-th component of the vector time series $\mathbf{y}_t^\nu$
%is the scalar time series $y_{\lambda,t}^\nu$, where $\nu=1,\dots,p$ and
%$\lambda=1,\dots,l$.
%
%However, the VAR model cannot be used with non-stationary variables. VECM ,
%developed by \cite{engle87}, is also a linear model for I(1) variables that are
%also cointegrated, see \cite{banerjee1993}. If cointegration exists, variable
%differences are stationary and they introduce an error correction term which
%adjusts coefficients to bring the variables back to equilibrium. 
%
%
%It is obtained re-writing equation (\ref{eq:var}) in terms of the new
%variable $\Delta\mathbf{y}_t=\mathbf{y}_t-\mathbf{y}_{t-1}$.
%The VECM model, expressed in terms those differences, takes the form:
%\begin{equation}\label{eq:vec}
%\Delta \mathbf{y}_t 
%= \boldsymbol{\Omega}\,\mathbf{y}_{t-1}
%  + \sum_{i=1}^{p-1} \boldsymbol{\Phi}_i^*\,\Delta\mathbf{y}_{t-i}
%  + \mathbf{c} + \boldsymbol{\epsilon}_t\,,
%\end{equation}
%\noindent
%where the coefficients matrices $\boldsymbol{\Phi}_i^*$ and 
%$\boldsymbol{\Omega}$, expressed in terms of the matrices
%$\boldsymbol{\Phi}_i$ of (\ref{eq:var}), are:
%\begin{align*}
%\boldsymbol{\Phi}_i^* 
%&:= -\sum_{j=i+1}^{p}\boldsymbol{\Phi}_j\,, \\
%\boldsymbol{\Omega}
%&:= -\left( \mathbb{I} - \boldsymbol{\Phi}_1 - \dots 
%    - \boldsymbol{\phi}_p \right)\,. 
%\end{align*}
%The following well known properties of the matrix $\boldsymbol{\Omega}$, see
%\cite{johansen1995}, will be useful in the sequel:
%\begin{itemize}
%\item
%If $\boldsymbol{\Omega} = \mathbf{0}$, there is no cointegration.
%\item 
%If $rank(\boldsymbol{\Omega})=l$, i.e., if $\boldsymbol{\Omega}$ has
%full rank, then the time series are not I(1) but stationary.
%\item
%If $rank(\boldsymbol{\Omega})=r$, $0<r<l$, then there is cointegration
%and the matrix $\boldsymbol{\Omega}$ can be expressed as
%$\boldsymbol{\Omega}=\boldsymbol{\alpha\beta}^\top$, where $\boldsymbol{\alpha}$
%and $\boldsymbol{\beta}$ are
%$l\times r$ matrices and
%$\text{rank}(\boldsymbol{\alpha})=\text{rank}(\boldsymbol{\beta})=r$.
%\item
%The columns of $\boldsymbol{\beta}$ contains the cointegration vectors and the rows of
%$\boldsymbol{\alpha}$ correspond with the adjusted vectors. 
%$\boldsymbol{\beta}$ is obtained by Johansen procedure,see \cite{johansen1988},
%whereas $\boldsymbol{\alpha}$ has to be determined as a variable in the VECM.
%\end{itemize}
%It is worth noticing that the factorization of the matrix
%$\boldsymbol\Omega$ is not unique, since for any $r \times r$
%non-singular matrix $\mathbf{H}$, $\boldsymbol{\alpha}^*:=\boldsymbol{\alpha}\mathbf{H}$,
%and $\boldsymbol{\beta}^*=\boldsymbol{\beta}(\mathbf{H}^{-1})^\top$ we have
%$\boldsymbol{\alpha\beta}^\top=\boldsymbol{\alpha}^*(\boldsymbol{\beta}^*)^\top$.
%If cointegration exists, then equation (\ref{eq:vec}) can be written
%as follows:
%\begin{equation}\label{eq:vecfull}
%\Delta\mathbf{y}_t 
%= \boldsymbol{\alpha\beta}^\top\mathbf{y}_{t-1} 
%  + \sum_{i=1}^{p-1}\boldsymbol{\Phi}_i^*\,\Delta\mathbf{y}_{t-i}
%  + \mathbf{c} + \boldsymbol{\epsilon}_t\,,
%\end{equation}
%\noindent
%which is a VAR model but for time series differences.
%
%
%Transposing each equation of the system (\ref{eq:vecfull}) we can write
%the VECM($p$) model in block-matrix form as:
%\begin{equation}\label{eq:vareq}
%\mathbf{B} = 
%\mathbf{A} \mathbf{X} + 
%\mathbf{E} \, , 
%\end{equation}
%%
%\noindent where $\mathbf{B}$ dimension is $((N-p)\times l)$, $\mathbf{A}$
%dimension is $((N-p)\times(r+(p-1)l +1))$, $\mathbf{X}$ dimension is $((r+(p-1)l
%+1)\times l)$ and $\mathbf{E}$ dimension is $((N-p)\times l)$:
%%
%\begin{alignat}{3}
%\mathbf{B}
%&= \begin{bmatrix}
%   \Delta\mathbf{y}_{p+1}^\top \\
%   \Delta\mathbf{y}_{p+2}^\top \\
%   \vdots \\
%   \Delta\mathbf{y}_N^\top
%   \end{bmatrix}
%&\quad
%\mathbf{X}
%&= \begin{bmatrix}
%   \boldsymbol{\alpha}^\top \\
%   \boldsymbol{\Phi}_1^{*\top} \\
%   \boldsymbol{\Phi}_2^{*\top} \\
%   \vdots \\
%   \boldsymbol{\Phi}_{p-1}^{*\top} \\
%   \mathbf{c}^\top
%   \end{bmatrix}
%&\quad
%\mathbf{E}
%&= \begin{bmatrix}
%   \boldsymbol{\epsilon}_{p+1}^\top \\
%   \boldsymbol{\epsilon}_{p+2}^\top \\
%   \vdots \\
%   \boldsymbol{\epsilon}_N^\top \\
%   \end{bmatrix}
%\end{alignat}
%\noindent and 
%\begin{align}
%\mathbf{A} 
%&= \begin{pmat}[{....|}]
%   \mathbf{y}_p^\top \boldsymbol{\beta} & \Delta \mathbf{y}_p^\top & \Delta\mathbf{y}_{p-1}^\top & \dots 
%                    & \Delta\mathbf{y}_2^\top & 1 \cr
%   \mathbf{y}_{p+1}^\top  \boldsymbol{\beta} &\Delta\mathbf{y}_{p+1}^\top & \Delta\mathbf{y}_p^\top & \dots
%                       & \Delta\mathbf{y}_3^\top & 1 \cr
%   \vdots & \vdots & \vdots & \ddots & \vdots & \vdots \cr
%   \mathbf{y}_{N-1}^\top  \boldsymbol{\beta} &\Delta\mathbf{y}_{N-1}^\top & \Delta\mathbf{y}_{N-2}^\top & \dots 
%                       & \Delta\mathbf{y}_{N-p-1}^\top & 1 \cr
%   \end{pmat}\, .
%\label{eq:Amatrix}
%\end{align}
%Taking into account the error term $\mathbf{E}$, equation~(\ref{eq:vareq}) 
%can be solved with respect to $\mathbf{X}$ using the OLS estimation.

\section{Methodology} \label{sec:51methodology}
\subsection{Motivation}

Cointegration vectors can be found applying the Johansen method which uses a
sample of the last historical data. However, VECM assumes cointegration vectors
do not change in time.  In fact, \cite{gregoryETal1996} addresses that the
long-run relationships between the time series  might change due to several
economic factors that can lead to structural breaks in the cointegration
relationship.  In order to show that the number of cointegration vectors depends
on the amount $L$ of historical data and the number of lags $p$ in the VECM, we
used a grid search.  We arbitrarily defined a grid of possible values for $L$
and $p$. $L$ goes throughout $[2,14]$ hours (1 $hour = 360$ data points) with a
step size of 4 hours and $p$ throughout $[1,5]$ with a step size of 1. The idea
was to show the variability of the number of cointegration vectors when we changed
these two parameters. We used four forex rates: EURUSD, GBPUSD, USDCHF and
USDJPY with 10-second frequency. Data started at 13:00 GMT of the 13th of August
2014, when the New York and London financial markets opened.

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
   \includegraphics[width=0.9\textwidth]{img/51_Fig1}
  \caption{Distribution of the number of cointegration vectors using $p=1$ lags.
  Four possible values for windows size $L$ are shown (2, 6, 10 and 14) hours (1
  hour = 360 data points).}
  \label{fig:hists}
\end{figure}

Figure \ref{fig:hists} shows the distribution of the number of cointegration
vectors given by the Johansen method for different values of $L=[2,6,10,14]$
hours and $p=1$ . This procedure was carried out by a sliding window of
historical data moving 1000 times. We observed that the distribution of
cointegration vectors changed with different values of $L$. When
$L=2$ hours, there was no cointegration in more than 60\% of the iterations.
Cointegration increased when $L=6$ hours and was maximum when $L=10$ hours where
one cointegration vectors was found for all 1000 iterations. The ocurrence of
cointegration started to decrease at $L=14$ hours.

From section \ref{sec:vecm} we know that $r=0$ means no cointegration and
$r=l$ (we are using four rates, so $l=4$) reveals that no process is I(1) but
stationary.  The interesting cases of cointegration are those where $r$ lies
strictly between $0$ and $4$, i.e. $0<r<4$.

@@

\subsection{Model comparison}
We compared our proposal, in terms of performance, with the naive forecast of
the random walk model and ARIMA. It is still difficult to outperform the random
walk model for standard econometric forecasting models  despite its
simplicity, see \cite{lo2011}. The random walk model is defined as:

@@

\subsection{Evaluation methods}

Forecast performance was evaluated using two different methods which are
frequently used in finance:
\begin{description}
\item[MSE]  Mean Square Error measures the distance between forecasts
and the true values and large deviations from the true value have a
large impact due to squaring forecast error.
\begin{equation}\label{eq:MSE}
\text{MSE} = 
\frac{\displaystyle \sum_{t=1}^{N} (\mathbf{y}_t-\hat{\mathbf{y}}_t)^2}{N}
\end{equation}
\item[$U$-statistic] the Theil's $U$-statistic, presented by
\cite{theil1966}, is a unit free measure obtained as the ratio between the root
MSE (RMSE) of a model and the RMSE of the naive random walk model. A
$U$-statistic less than 1 implies the performance is better than the naive
model.
\end{description}

\section{Experimental results}
\label{sec:51results}
\subsection{Data}
All the experiments and AVECM tests were carried out using four foreign exchange rates all related to
USD: EURUSD, GBPUSD, USDCHF and USDJPY. We chose the most traded rates related
to USD so they were likely to be cointegrated. 
This data was collected from
\cite{Dukascopy2014}, a free database which gives access to the Swiss Foreign
Exchange marketplace.

@@

\subsection{Unit root tests}

@@

\subsection{Performance accuracy}

@@

\subsection{Parallel implementation}

@@

\section{Conclusions}
\label{sec:51conclusions}
Cointegration in financial time series has been largely studied and the Johansen
method is commonly used to obtain cointegration relationships. In practice, it
has been found that cointegration relations change with time. However,
model-based cointegration such as VECM assumes that cointegration remains
unchanged in time. We empirically showed that the Johansen method is sensitive
to the number of lags but also to the amount of data considered.
